\documentclass[15pt]{beamer}

\usepackage[utf8]{inputenc}
\usepackage{xcolor}
\usepackage{hyperref}

\usetheme{Madrid}
\usecolortheme{beaver}

\title{Computer Networks Laboratory}
\subtitle{Chapter 0: Introduction}
\author{Ahmad Khonsari}
\institute{
    ECE Department\\
    Tehran University \newline
    \url{http://ece.ut.ac.ir/portal/}
}
\date{Spring 1399}
\subject{Computer Science}

\setbeamercolor{section number projected}{bg=red,fg=white}
\setbeamercolor*{item}{fg=red}

\begin{document}

\frame{\titlepage}

\begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents
\end{frame}

\section{Teaching Staff}
\begin{frame}
\frametitle{Teaching Staff}
\begin{itemize}
    \item Instructor:
    \begin{itemize}
        \item Ahmad Khonsari
    \end{itemize}
    \item Chief Teaching assistants:
    \begin{itemize}
        \item Amir Haji Ali Khamseh’i: khamse@ut.ac.ir
    \end{itemize}
    \item Teaching assistants:
    \begin{itemize}
        \item Mohammad Borhani: borhani.m@ut.ac.ir
    \end{itemize}
    \item Contact time: Tuesday-Wenesday (netlab-softlab (azmayeshagah narmafzar))
\end{itemize}

\end{frame}

\section{Course Goals}
\begin{frame}
    \frametitle{Course Goals}

    \begin{itemize}
        \item Learn how to config host and networking device
        \item Learn how to network device work
        \item Learn how to simulate network
        \item Learn new network protocol (SNMP, \dots)
        \item IPv4, IPv6, Routing, ARP, RIP, OSPF, UDP, TCP, Multicast, NAT, HTTP, DHCP, SNMP, Security, ...         
        \item During the study you need to read extra book and documents for do your lab (has not been taught in the Network lesson)
    \end{itemize}

\end{frame}

\section{Course Background}
\begin{frame}
    \frametitle{Course Background}

    \begin{itemize}
        \item Computer Net Lab is a one unit course and it is a follow up of the Computer Networks course.
        \item The materials you have learned in the computer network course is necessary but not sufficient to do Computer Net Lab.
              You will learn at least 40\% of networking materials that you didn't see in the main course (such as IPV6, Spanning Tree, Multicast, MPLS) . It is necessary to learn the necessary background before attending each session.
        \item The grading of this course is divided to 3 main parts.
        \begin{enumerate}
            \item Reading the background material of each lab before attending each session. There is a quiz in the beginning of each lab of the main materials of the current session.
            \item Carrying out the lab within the time slot that assigned for each lab and prepare the report
            \item A final exam. In the current semester due to Covid 19, we may ask students to voluntarily prepare and fulfil an extra lab instead of exam.
        \end{enumerate}
    \end{itemize}


\end{frame}

\section{Course Format}
\begin{frame}
    \frametitle{Course Format}

    \begin{itemize}
        \item 10 sessions
        \item Cover the “TCP/IP essential: A lab based approach” + one new Lab
        \item 10 lab reports + TA quizzes
        \item {\color{red} Don’t accept delay (must submit and finish your work in time)}
    \end{itemize}

\end{frame}

\section{Grading}
\begin{frame}
    \frametitle{Grading}

    \begin{itemize}
        \item For remote class when COVID-19
        \begin{itemize}
            \item {\color{red} 1.8} for each lab report
        \end{itemize}
        \item For normal class
        \begin{itemize}
            \item {\color{red} 1.4} for finishing up the lab
            \item {\color{red} 0.4} for each lab report (1 week)
        \end{itemize}
        \item {\color{red} 0.2} to {\color{red} 0.3} for each quiz (0.1 Additional Score)
        \item Don’t submit image, except for report questions that specifically ask for an image.
        \item 
        \item You can only object to your grades within {\color{red} 2 days} of receiving them. We do not consider objections after that. 
        
    \end{itemize}

\end{frame}

\section{Course Materials}
\begin{frame}
    \frametitle{Course Materials}

    \begin{itemize}
        \item Textbook: TCP/IP Essential, a lab based approach
        \item \url{https://github.com/UT-Network-Lab/TCP-IP-Laboratory}
        \item GNS3 on Linux (physical or virtual machine)
        \item Cisco router image (c3725-adventerprisek9-mz.124-25d.image)
        \item Docker images (netlab/term, netlab/gui)
        \item GNS3 Portable projects (Figures)
    \end{itemize}

\end{frame}

\section{Class rules}
\begin{frame}
    \frametitle{Class rules}

    LATE ARRIVAL:
    \begin{itemize}
        \item below 15 min $\rightarrow$ don’t worry (but you may miss quiz)
        \item {\color{orange} between 15 and 30 min $\rightarrow$ get -0.5 for your session grade}
        \item {\color{red} more than 30 min $\rightarrow$ don’t enter the class}
    \end{itemize}

\end{frame}

\end{document}